---
title: "Homework 4"
author: "Matthew Wynne"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  2,3 (Pg. 151); 2,4 (Pg. 156); 1,2 (Pgs. 160-161); 2 (Pg. 163); 2,3,4 (Pg. 168)

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
```

# Chapter 9 (Pg. 151)

```{r}
table2
```

##  Exercise 2
```{r}
cases2 = table2 %>%
  filter(type == "cases")

pop2 = table2 %>%
  filter(type == "population")

table2_cases_per_cap = tibble(
  country = cases2$country,
  year = cases2$year,
  cases = cases2$count,
  population = pop2$count,
  cases_per_capita = (cases / population) * 10000)

table2_cases_per_cap
```
```{r}
table4 = tibble(
  country = table4a$country,
  `1999` = table4a$`1999` / table4b$`1999`,
  `2000` = table4a$`2000` / table4b$`2000`)

table4
```
`table4a` and `table4b` were easier to work with since you could find the rate in one step, while `table2` was harder because you had to split up the count variable into both cases and population and then you could find the rate.

##  Exercise 3
```{r}
ggplot(table2_cases_per_cap, aes(year, cases, color = country)) +
  geom_point() +
  geom_line()
```
In order to do plot this with `table2`, you had to make cases its own column in the data frame, which we previously did in exercise 2.

# Chapter 9 (Pg. 156)

##  Exercise 2
```{r}
table4a %>%
  gather(`1999`, `2000`, key = "year", value = "cases")
```
There needs to be back ticks (``) around the years since they are column names, otherwise it will try to select the 1999th and 2000th columns.

##  Exercise 4
```{r}
preg <- tribble(
~pregnant, ~male, ~female, "yes", NA, 10, "no", 20, 12
)
preg
```
```{r}
gather(preg, key = gender, value =  value, male:female, na.rm = TRUE)
```
I used `gather()` to create a new column, `gender`, and remove the missing value for males. The variables are now `pregnant`, `gender`, and `value`.

# Chapter 9 (Pgs. 160-161)

##  Exercise 1
```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>%
  separate(x, c("one", "two", "three"))
```
For this one, the `extra` argument in `separate()` gives a warning that too many pieces were given, and automatically drops the "g".

```{r}
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>%
  separate(x, c("one", "two", "three"))
```
For this one, the `fill` argument in `separate()` gives a warning that not enough pieces were given, and automatically it in with an "NA".

##  Exercise 2
```{r}
tibble(x = c("a,b,c", "d,e,f", "g,h,i")) %>%
  separate(x, c("one", "two", "three"), remove = FALSE)
```
The `remove` argument in `unite()` and `separate()` is automatically set to true and removes the input column. In the example above, when `remove` is set to `FALSE`, the input column, `x`, remains in the data frame.

# Chapter 9 (Pg. 163)

##  Exercise 2
```{r}
tibble(x = c("a,b,c", "d", "g,h,i")) %>%
  separate(x, c("one", "two", "three"), fill = "left")
```
```{r}
tibble(x = c("a,b,c", "d", "g,h,i")) %>%
  separate(x, c("one", "two", "three"), fill = "right")
```
As you can see above, when `fill = "left"`, the NAs are added to the left-most columns. When `fill = "right"`, the NAs are added to the right-most columns.

# Chapter 9 (Pg. 168)

##  Exercise 2
```{r}
who2 = who %>%
  gather(code, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>%
  mutate(
        code = stringr::str_replace(code, "newrel", "new_rel")
      ) %>%
  separate(code, c("new", "var", "sexage")) %>%
  select(-new, -iso2, -iso3) %>%
  separate(sexage, c("sex", "age"), sep = 1)

who2
```
```{r}
who %>%
  gather(code, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>%
  separate(code, c("new", "var", "sexage")) %>%
  select(-new, -iso2, -iso3) %>%
  separate(sexage, c("sex", "age"), sep = 1)
```
When we neglect the `mutate()` step, we are given an error from `separate()` that we have too few values.

```{r}
who %>%
  gather(code, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>%
  separate(code, c("new", "var", "sexage")) %>%
  filter(new == "newrel") %>%
  head(5)
```
Looking at the above code, you can see that when we skip the `mutate()` step, `sexage` is completely empty, and all of it's data is contained in `var`.

##  Exercise 3
```{r}
who %>%
  select(country, iso2, iso3) %>%
  distinct()
```
As you can see, all of the above are just different representations of the same countries in the data set.

##  Exercise 4
```{r}
who2 %>%
  group_by(country, year, sex) %>%
  filter(year > 1995) %>%
  summarize(cases = sum(value)) %>%
  unite(country_sex, country, sex, remove = FALSE) %>%
  ggplot(aes(year, cases, group = country_sex, color = sex)) +
    geom_line()
```
