---
title: "Homework 5"
author: "FIRSTNAME LASTNAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  1,3 (Pg. 227); 2 (Pg. 232); 1,2 (Pg. 235); 

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
```

# Chapter 12 (Pg. 227)

##  Exercise 1
```{r}
ggplot(gss_cat, aes(rincome)) +
  geom_bar()
```
```{r}
ggplot(gss_cat, aes(rincome)) +
  geom_bar() +
  coord_flip()
```

The first chart is hard to understand because all of the ranges on the x-axis are overlapping. This is easily fixed by flipping the axes, as done above.

##  Exercise 3
```{r}
unique(select(gss_cat, denom))

ggplot(gss_cat, aes(denom)) +
  geom_bar() +
  coord_flip()
```

`denom` applies to Christianity. By looking at the column's unique values above, I was about to determine this. You could also do a horizontal bar chart to see all of the unique denominations.

# Chapter 12 (Pg. 232)

##  Exercise 2
```{r}
levels(gss_cat$marital)
```
Arbitrary

```{r}
levels(gss_cat$race)
```
Arbitrary

```{r}
levels(gss_cat$rincome)
```
Principled

```{r}
levels(gss_cat$partyid)
```
Principled within parties, but arbitrary between parties.

```{r}
levels(gss_cat$relig)
```
Arbitrary

```{r}
levels(gss_cat$denom)
```
Arbitrary

# Chapter 12 (Pg. 235)

##  Exercise 1
```{r}
gss_cat %>%
  mutate(
    party = fct_collapse(partyid,
      Republican = c("Strong republican", "Not str republican"),
      Democrat = c("Not str democrat", "Strong democrat"),
      Independent = c("Ind,near rep", "Independent", "Ind,near dem"),
      Other = c("No answer", "Don't know", "Other party")
      )
    ) %>%
  group_by(year, party) %>%
  summarize(
    n = n(),
  ) %>%
  ggplot() +
  geom_smooth(aes(year, n, color = party), se = FALSE)
```

##  Exercise 2

To collapse `rincome` into a smaller set of categories, you could consolidate “No answer”, “Don’t know”, and “Refused” into a single category. You could also change the income ranges to $5000 increments versus having smaller increments at the lower ranges.