---
title: "Homework 1"
author: "MATTHEW WYNNE"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:** 1-5 (Pgs. 6-7); 1-2, 5 (Pg. 12); 1-5 (Pgs. 20-21); Open Response

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Chapter 1 (Pgs. 6-7)

##  Exercise 1
```{r}
ggplot(data=mpg)
```

I see absolutely nothing. There is just a blank space for a graph. Why am I even doing this nonsense?

## Exercise 2
```{r}
dim(mpg)
nrow(mpg)
ncol(mpg)
```
There are 234 rows and 11 columns in the dataset **mpg**.

##  Exercise 3
```{r, eval=FALSE}
?mpg
```

```{r}
unique(mpg$drv)
```


The variable **drg** is a factor variable that takes the following values:

- "f" = front-wheel drive
- "r" = rear-wheel drive
- "4" = 4-wheel drive

##  Excercise 4
```{r}
ggplot(data=mpg,aes(x=hwy,y=cyl)) +
  geom_point() + 
  xlab("Highway Miles Per Gallon") +
  ylab("Number of Cylinders")
```


##  Excercise 5
```{r}
ggplot(data=mpg,aes(x=class,y=drv)) + 
  geom_point() + 
  xlab("Type of Car") +
  ylab("Type of Drive")
```

Scatter plots are not meant to visualize the relationship between two categorical/qualitative variables. 

# Chapter 1 (Pg. 12)

##  Exercise 1

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

The code above does not put the aesthetic `"color = "blue"` outside of the `aes()` function, which is necessary to set an aesthetic manually. The correct method is shown below.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

##  Exercise 2

As seen in **mpg** below, the categorical variables are listed with  `<chr>` underneath them. The continuous variables are listed with `<int>` or `<dbl>` underneath them.
```{r}
mpg
```

The categorical variables for **mpg** are the following:
  
- "manufacturer"
- "model"
- "trans"
- "drv"
- "fl"
- "class"

The continuous variables for **mpg** are the following:

- "displ"
- "year"
- "cyl"
- "cty"
- "hwy"

##  Exercise 5

For shapes that have a border, the `stroke` aesthetic controls the size of the border.

# Chapter 1 (Pgs. 20-21)

##  Exercise 1

- **line chart**: `geom_line()`, `geom_path()`, or `geom_step()`
- **boxplot**: `geom_boxplot()`
- **histogram**: `geom_histogram()`
- **area chart**: `geom_area()`

##  Exercise 2

Since there is a mapping in the ggplot() argument and two geoms are given, it will be treated as a global mapping with `color = drv`. `geom_point()` will indicate drive train by the color of the points and `geom_smooth()`  will indicate drive train by the color of the lines.

```{r}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = drv)
) +
  geom_point() +
  geom_smooth(se = FALSE)
```

As you can see in the graph above, the prediction was correct.

##  Exercise 3

`show.legend = FALSE` is used to remove a legend from a graph that would normally generate one. If it is removed, the legend is still generated. This indicates that the legend is printed by default unless otherwise indicated. This was used earlier in the chapter because we were trying to demonstrate that the `group` aesthetic does not add distinguishing features to the different objects, and therefore does not need a legend.

##  Exercise 4

The `se` argument for `geom_smooth()` controls if the confidence interval around the line is shown or not. It is set to `TRUE` by default. The default level of confidence is 0.95 and can be adjusted using the `level` argument.

##  Exercise 5

I don't know if they will look different. Let me check.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

They do not look different. I am incredibly surprised.

# Open Response

For this exercise, use the **diamonds** dataset in the tidyverse. Use `?diamonds` to get more information about the dataset.

## Step 1: Select 1 numeric variable and 2 categorical variables. Create a graphic using `geom_boxplot()` and `facet_wrap` to illustrate the empirical distributions of the sample.

```{r}
ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = carat, y = clarity)) +
  facet_wrap(~ color)
```

## Step 2: Choose 2 numeric variables and 2 categorical variables and creatively illustrate the relationship between all the variables. 
```{r}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price , color = clarity)) +
  facet_grid(~cut)
```